\PaperTitle

### 2020-01-25 paper/filter_bubbles.tex
typos

Risk aversion captures the how different users react to the risk associated to a particular consumption opportunity. It is formalized as follows: Suppose for a given gamble $x$ taking real values and following distribution $p$, there is a certain amount of money that leaves the user indifferent between taking the gamble and taking this sure amount, call it $\delta(x)$, termed the \textit{certainty equivalent} of $x$. A user $i$ is more risk-averse than another user$j$ when for every gamble $x$, user $i$ requires a lower value to give up the gamble when compared to user $j$, i.e. $\delta_i(x)$. Therefore, a more risk-averse user is more willing to avoid the risk of taking the gamble. We assume that the utility function takes a flexible functional form $u(x)=1-\exp(-\gamma x)$ for $\gamma\ne0$ and $u(x)=x$ for $\gamma=0$ -- known as constant absolute risk-aversion preferences (from hereon CARA). Higher $\gamma$ implies higher risk-aversion with $\gamma=0$ corresponding to the risk-neutral case and $\gamma>0$ to the risk-averse one. Our formulations here follow standard economic decision-theory (see \cite{mas1995microeconomic} for a textbook treatment of these topics).
\par

@@

We denote by $x_{i,n}$ user $i$'s realized value from consuming item $n$. In particular, we consider that the realized value derived from a given item can be decomposed in the following manner: $x_{i,n}= v_{i,n} + \beta v_n$, where $v_{i,n}$ denotes an idiosyncratic component -- i.e. user $i$'s idiosyncratic taste for product $n$ --  and $v_{n}$, a common-value component. One can interpret $v_n$ as a measure of how much item $n$ is valued in society in general and, in a sense, $v_{i,n}$ denotes how $i$ diverges from this overall ranking. The scalar $\beta \in \mathbb{R}_{t}$ denotes the degree to which valuations are idiosyncratic to each user or common across users. If $\beta=0$, it is impossible to generate meaningful predictions of any one's individual preferences based on others, while if $\beta$ is large, every individual has similar preferences.
\par
Stacking values in vector-form, we get the vector of values associated with each item 
$${\left(x_{i,n}\right)}_{n \in \mathcal{J}}=:X_i =V_i+ \beta V $$
where $V_i ={\left(v_{i,n}\right)}_{n \in \mathcal{J}}$ and $V={\left(v_{n}\right)}_{n \in \mathcal{J}}$.
\par
\xhdr{User Decision-Making.}
We assume the user makes $T$ choices and therefore can only consume up to $T$ items, where $T$ is but a small fraction of $N$. This captures the idea that users are faced with an immense choice set but that ultimately they end up experiencing (and learning) about just a small fraction of it. Since this is a sequential decision-making problem under uncertainty, in principle users face an exploration-exploitation trade-off. However, for tractability, we impose that users are myopic and every period consume the product that they have not yet tried ($n_i^t$) that gives them the highest expected utility given the information from past consumption offers ($C_i^{t-1}=(n_i^1,...,n_i^{t-1})$) and their initial beliefs.
\par
\xhdr{User Beliefs.} We assume that all the true realized utilities are drawn at $t = 0$. However, users do not know the realized utilities before consuming an item, but rather have beliefs over the realized utilities.
Formally, user $i$ starts with some beliefs about $X_i$, namely that the idiosyncratic and common-value parts of the valuations are independent -- $V_i \perp \!\!\! \perp V$ -- and that each is multivariate normal 
\begin{enumerate}
\item $V_i \sim \mathcal N (\overline V_i, \Sigma_i)$ 
\item $V \sim \mathcal N(\overline V, \Sigma)$ with $\overline V =0$
\end{enumerate}
We impose the normality assumption for two reasons. The first is that this allows for simple and tractable belief updating. The second is that it allows us to incorporate an easily interpretable correlation structure between the items.
\par
Keeping with the assumption that $V_i$ represents idiosyncratic deviations from $V$, we assume that, on the population level, prior beliefs $\overline V_i=\left(\overline v_{i,n}\right)_{n \in \mathcal{J}}$ are drawn independently from a jointly normal distribution, where $\overline v_{i,n} \sim \mathcal N (0, \overline \sigma^2)$ are independent and identically distributed. These $\overline v_{i,n}$ denote the prior belief that individual $i$ holds about the her valuation over item $n$. As people are exposed to different backgrounds, their beliefs about what is good for them also varies and $\overline v_{i,n}$ denotes this idiosyncrasy at the level of prior beliefs.
\par

@@

\subsection{User Homogenization}
In this section we focus on across user comparisons and investigate how the consumed set of items across users varies across different recommendation regimes and parameter values. In particular we look at the degree of \textit{homogenization} between users. Similar to other papers that study the degree of homogenization in recommender systems (e.g. \cite{chaney2018algorithmic}) we measure homogeneity via the Jaccard index between the consumption sets of users:
\begin{align*}
H:=\frac{1}{|I|(|I|-1)}\sum_{i,j \in I: i \ne j}d_J(C_i^T,C_j^T)
\end{align*}
where $d_J$ denotes the Jaccard index and $H \in [0,1]$.
\begin{finding}\label{finding_homogeneity}
The impact of recommendation on homogeneity is as follows:
\begin{enumerate}
\item Highest under recommendation and lowest under no recommendation;
\item Increasing in $\beta$, or the weight of the common-value component;
\item Decreasing in $\rho$ for partial recommendation, but weakly increasing in $\rho$ for no recommendation.
\end{enumerate}
\end{finding}
Finding~\ref{finding_homogeneity} summarizes our findings on the impact of recommendation on user homogeneity. First, we study how the degree of homogenization varies as we increase $\beta$, the weight of the common value component. As $\beta$ increases we expect that users should become increasingly homogeneous as the realized utilities of the items are now increasingly similar. Figure \ref{fig:beta_homo} confirms that as the weight of the common-value component, $\beta$, increases users consume increasingly similar sets of items. The homogenization effect is strongest under the recommendation regime since the revelation of the common-value component induces users to consume products in similar areas of the product space. As $\beta$ increases, some amount of homogenization is optimal as can be seen from the oracle case. However, since users in the no recommendation regime do not know the common-value component they engage in local consumption in different areas of the product space which leads to less than optimal homogeneity.
\par
We next study how the degree of homogeneity varies as $\rho$ increases. Figure \ref{fig:cor_homo} shows how homogeneity varies as $\rho$ increases across the different recommendation regimes. The most striking finding is that homogeneity decreases as $\rho$ increases in the recommendation regime. As was highlighted in Findings \ref{finding_local_consumption} and \ref{finding_diversity}, the degree of local consumption increases with $\rho$. Even though the revelation of the common-value component induces them to search in similar parts of the product space, their idiosyncratic components induce them to consume products in a more localized area of the product space as $\rho$ increases which leads to a decline in homogeneity.

>>>>

Risk aversion captures the how different users react to the risk associated to a particular consumption opportunity. It is formalized as follows: a given gamble $x$ takes real values and follows distribution $p$. Then, for every gamble $x$, there is a certain amount of money that makes the user indifferent between taking the gamble or taking the sure amount of money. This sure amount of money is called the \textit{certainty equivalent} of gamble $x$ and is denoted as $\delta(x)$. A user $i$ is more risk-averse than another user $j$ when for every gamble $x$, user $i$ requires a lower value to give up the gamble when compared to user $j$, i.e. $\delta_i(x)$. Therefore, a more risk-averse user is more willing to avoid the risk of taking the gamble. We assume that the utility function takes a flexible functional form $u(x)=1-\exp(-\gamma x)$ for $\gamma\ne0$ and $u(x)=x$ for $\gamma=0$ -- known as constant absolute risk-aversion preferences (from hereon CARA). Higher $\gamma$ implies higher risk-aversion with $\gamma=0$ corresponding to the risk-neutral case and $\gamma>0$ to the risk-averse one. Our formulations here follow standard economic decision-theory (see \cite{mas1995microeconomic} for a textbook treatment of these topics).
\par

@@

We denote by $x_{i,n}$ user $i$'s realized value from consuming item $n$. In particular, we consider that the realized value derived from a given item can be decomposed in the following manner: $x_{i,n}= v_{i,n} + \beta v_n$, where $v_{i,n}$ denotes an idiosyncratic component -- i.e. user $i$'s idiosyncratic taste for product $n$ --  and $v_{n}$, a common-value component. One can interpret $v_n$ as a measure of how much item $n$ is valued in society in general and, in a sense, $v_{i,n}$ denotes how $i$ diverges from this overall ranking. The scalar $\beta \in \mathbb{R}_{t}$ denotes the degree to which valuations are idiosyncratic to each user or common across users. If $\beta=0$, it is impossible to generate meaningful predictions of any one's individual preferences based on others, while if $\beta$ is large, every individual has similar preferences.
\par
Stacking values in vector-form, we get the vector of values associated with each item 
$${\left(x_{i,n}\right)}_{n \in \mathcal{J}}=:X_i =V_i+ \beta V $$
where $V_i ={\left(v_{i,n}\right)}_{n \in \mathcal{J}}$ and $V={\left(v_{n}\right)}_{n \in \mathcal{J}}$.
\par
\xhdr{User Decision-Making.}
We assume the user makes $T$ choices and therefore can only consume up to $T$ items, where $T$ is a small fraction of $N$. This captures the idea that users are faced with an immense choice set but that ultimately they end up experiencing (and learning) about just a small fraction of it. Since this is a sequential decision-making problem under uncertainty, in principle users face an exploration-exploitation trade-off. However, for tractability, we impose that users are myopic and every period consume the product that they have not yet tried ($n_i^t$) that gives them the highest expected utility given the information from past consumption offers ($C_i^{t-1}=(n_i^1,...,n_i^{t-1})$) and their initial beliefs.
\par
\xhdr{User Beliefs.} We assume that all the true realized utilities are drawn at $t = 0$. However, users do not know the realized utilities before consuming an item, but rather have beliefs over the realized utilities.
Formally, user $i$ starts with some beliefs about $X_i$, namely that the idiosyncratic and common-value parts of the valuations are independent -- $V_i \perp \!\!\! \perp V$ -- and that each is multivariate normal 
\begin{enumerate}
\item $V_i \sim \mathcal N (\overline V_i, \Sigma_i)$ 
\item $V \sim \mathcal N(\overline V, \Sigma)$ with $\overline V =0$
\end{enumerate}
We impose the normality assumption for two reasons. The first is that this allows for simple and tractable belief updating. The second is that it allows us to incorporate an easily interpretable correlation structure between the items. The precise formulation of $\Sigma$ and $\Sigma_i$ that we consider is defined below when we discuss user learning.
\par
Keeping with the assumption that $V_i$ represents idiosyncratic deviations from $V$, we assume that, on the population level, prior beliefs $\overline V_i=\left(\overline v_{i,n}\right)_{n \in \mathcal{J}}$ are drawn independently from a jointly normal distribution, where $\overline v_{i,n} \sim \mathcal N (0, \overline \sigma^2)$ are independent and identically distributed. These $\overline v_{i,n}$ denote the prior belief that individual $i$ holds about the her valuation over item $n$. As people are exposed to different backgrounds, their beliefs about what is good for them also varies and $\overline v_{i,n}$ denotes this idiosyncrasy at the level of prior beliefs.
\par

@@

\subsection{User Homogenization}
In this section, we focus on comparisons across users and investigate how the consumed set of items across users varies across different recommendation regimes and parameter values. In particular we look at the degree of \textit{homogenization} between users. Similar to other papers that study the degree of homogenization in recommender systems (e.g. \cite{chaney2018algorithmic}) we measure homogeneity via the Jaccard index between the consumption sets of users:
\begin{align*}
H:=\frac{1}{|I|(|I|-1)}\sum_{i,j \in I: i \ne j}d_J(C_i^T,C_j^T)
\end{align*}
where $d_J$ denotes the Jaccard index and $H \in [0,1]$.
\begin{finding}\label{finding_homogeneity}
The impact of recommendation on homogeneity is as follows:
\begin{enumerate}
\item Highest under recommendation and lowest under no recommendation;
\item Increasing in $\beta$, or the weight of the common-value component;
\item Decreasing in $\rho$ for partial recommendation, but weakly increasing in $\rho$ for no recommendation.
\end{enumerate}
\end{finding}
Finding~\ref{finding_homogeneity} summarizes our findings on the impact of recommendation on user homogeneity. First, we study how the degree of homogenization varies as we increase $\beta$, the weight of the common value component. As $\beta$ increases we expect that users should become increasingly homogeneous as the realized utilities of the items are now increasingly similar. Figure \ref{fig:beta_homo} confirms that as the weight of the common-value component, $\beta$, increases users consume increasingly similar sets of items. The homogenization effect is strongest under the recommendation regime since the revelation of the common-value component induces users to consume products in similar areas of the product space. As $\beta$ increases, some amount of homogenization is optimal as can be seen from the oracle case. However, since users in the no recommendation regime do not know the common-value component they engage in local consumption in different areas of the product space which leads to less than optimal homogeneity.
\par
We next study how the degree of homogeneity varies as $\rho$ increases. Figure \ref{fig:cor_homo} shows how homogeneity varies as $\rho$ increases across the different recommendation regimes. The most striking finding is that homogeneity decreases as $\rho$ increases in the recommendation regime. As was highlighted in Findings \ref{finding_local_consumption} and \ref{finding_diversity}, the degree of local consumption increases with $\rho$. Even though the revelation of the common-value component induces them to search in similar parts of the product space, their idiosyncratic components induce them to consume products in a more localized area of the product space as $\rho$ increases which leads to a decline in homogeneity.
